---
title: "nacleaner guide"
output: html_document
date: "2026-05-03"
author: "Mason Good"
---

# About This Package

I built `nacleaner` as my final project for LIS 4370 R Programming. The
package handles missing values in data frames.
There are bigger and more capable packages out there (mice or naniar) that I took inspiration from but wanted to try to build my own. 

# The Dataset

For testing I used the built-in `airquality` dataset which includes na values. 

```{r}
library(nacleaner)
summary(airquality)
```

# Function 1: na_report()

The first thing you usually want to know is where the missing values
are. `na_report()` returns a small data frame showing the count and
percentage for each column.

```{r}
na_report(airquality)
```


# Function 2: drop_na_rows()

Sometimes you just want clean rows. `drop_na_rows()` is a thin wrapper
around `complete.cases()` but with input validation.

```{r}
nrow(airquality)
nrow(drop_na_rows(airquality))
```

You lose 42 rows from airquality this way, which is a lot. That is why
the next function exists.

# Function 3: fill_na()

Instead of dropping rows, you can fill them. The function only touches
numeric columns and gives you a choice between mean and median.

```{r}
filled <- fill_na(airquality, method = "median")
na_report(filled)
```

I used median in the example because Ozone is right-skewed and the
mean would be pulled up by high readings.

# What I Would Add Next

If I keep working on this I would add:

- A method that imputes by group (fill within each Month for example)
- Visualization of the NA pattern
- Handling of character columns

# AI Disclosure

I used Claude to help debug my package. For whatever reason my description file kept giving me lots of issues. Claude helped me generate a description file that the devtools::check() liked. 
